---
title: "Mini-Project 2"
author: "Sneha, Pema, Tenzin"
date: "October 31, 2017"
output: html_document
---

##The Incumbency Advantage

**What is the Incumbent Advantage?**
 
As the incumbent, one holds many advantages over the challenger. Some of these advantages come in the form of:
Funding
Name recognition
National attention
Fundraising and Campaign Bases
Control over the instruments of govt.
Successful Campaign Experience
Presumption of success and voters

For our project we are looking at whether incumbents held a monetary advantage over their challengers in the U.S. elections for 2012. In order to do this we will take a look at monetary contributions by committees to challengers, incumbents, and open office. 

```{r, include=FALSE, message=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
```
```{r eval=FALSE}
head(house_elections)
```

```{r, eval = FALSE}
head(committees)
```
```{r, eval = FALSE}
head(candidates)
```
```{r, eval = FALSE}
head(contributions)
```

```{r}
candidate_contributions <- candidates %>%
  #joining candidates and contributions by "cand_id"
  inner_join(contributions, by = "cand_id") %>%
  #selecting the relevant variables
  select(cand_id, cand_name, cand_ici, cand_election_yr, cand_party_affiliation, cmte_id, transaction_type, name, transaction_dt, transaction_amt)
```

```{r}
committee_to_cand <- candidate_contributions %>%
  #joining candidates, committees and contributions
  inner_join(committees, by = "cmte_id") %>%
  #select relevant columns
  select(cmte_id, cmte_name, cmte_party_affiliation, transaction_type, name, transaction_dt, transaction_amt, cand_name, cand_party_affiliation, cand_ici)

#view dimensions of the new tbl_dr
dim(committee_to_cand)
```
```{r}
#set the desired values of transaction_type
#transaction <- c("24A", "24E")
transaction <- c("24E") #look for transactions advocating election of a candidate
committee_to_cand <- committee_to_cand %>%
  #filter transaction_type to values in 'transaction' i.e either "24A" or "24E"
  filter(transaction_type %in% transaction) %>%
  mutate(transaction_dt = mdy(transaction_dt))

#view dimensions to ensure there is enough data to use in our anaysis
dim(committee_to_cand)
```

```{r}
#filter date to year 2012 i.e transactions between Jan 1st, 2012 to Dec 31st, 2012
committee_trans_2012 <- committee_to_cand %>%
  filter(transaction_dt >= "2012-01-01" & transaction_dt <= "2012-12-31")
head(committee_trans_2012)
dim(committee_trans_2012)
```

```{r}
susanb_cand_fund <- committee_trans_2012 %>%
  #filtered for a certain committee
  filter(cmte_name == "SUSAN B ANTHONY LIST INC. CANDIDATE FUND") %>%
  #grouped by whether candidate is incumbent, challenger or 
  group_by(cmte_name, cand_ici) %>%
  summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))
head(susanb_cand_fund)
```

```{r}
ggplot(susanb_cand_fund, aes(x = cmte_name, y = cont_to_ici,
                             fill = cand_ici))+
  coord_flip()+
  geom_bar(stat = "identity") + 
  scale_x_discrete("Committee", labels = c("Susan Anthony")) +
  scale_y_continuous("Total Contributions", limits = c(0,12000), breaks = seq(0,12500, 6000), labels = c("$0", "$6000", "$12000"))+
  scale_fill_manual(values = c("#999999", "#E69F00", "#56B4E9"),
                    name = "Candidate\nStatus",
                    breaks = c("C", "I", "O"),
                    labels = c("Challenger", "Incubent", "Open Office"))
#  guides(fill = guide_legend("Candidate
#Status"))
```

```{r}
#tidy
susanb_cand_fund_spread <- susanb_cand_fund %>%
  spread(key = "cand_ici", value = "cont_to_ici") %>%
  mutate(total_contributions = C+I+O)
head(susanb_cand_fund_spread)
```

```


```{r}
#tidy
susanb_cand_fund_spread <- susanb_cand_fund %>%
  spread(key = "cand_ici", value = "cont_to_ici") %>%
  mutate(total_contributions = C+I+O)
head(susanb_cand_fund_spread)
```








