---
title: "Mini-Project 2"
author: "Sneha, Pema, Tenzin"
date: "October 31, 2017"
output: html_document
---

##The Incumbent Advantage


Before we begin let us define some basic terms:

**Who is the ‘incumbent’?**

*The ‘incumbent’ is the current holder of an office or a political post.*

**Who is the ‘challenger’?**

*The ‘challenger’ is the person who wants to hold the office or political post next, i.e the person challenging the incumbent.* 

*For example, in the 2012 U.S. presidential elections, Mitt Romney was challenger to Barack Obama’s incumbency.*

**What is an 'Open Seat?**

*An election with no incumbent runner is called an open seat.*


**What is the Incumbent Advantage?**
 
As the incumbent, one holds many advantages over the challenger. Some of these advantages come in the form of:

- Funding

- Name recognition

- National attention

- Fundraising and Campaign Bases

- Control over the instruments of govt.

- Successful Campaign Experience

- Presumption of success and voters

For our project we are looking at whether incumbents held a monetary advantage over their challengers in the U.S. elections for 2012. In order to do this we will take a look at monetary contributions by committees to challengers, incumbents, and open office by state. More specifically we will be looking at transactions of type '24E', i.e. transactions made to advocate the election of a candidate.

```{r, include=FALSE, message=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
```
```{r eval=FALSE}
head(house_elections)
head(committees)
head(contributions)
```


```{r}
candidate_contributions <- candidates %>%
  #joining candidates and contributions by "cand_id"
  inner_join(contributions, by = "cand_id") %>%
  #selecting the relevant variables
  select(cand_id, cand_name, cand_ici, cand_election_yr, cand_party_affiliation, cmte_id, transaction_type, name, transaction_dt, transaction_amt, state)
```

```{r}
committee_to_cand <- candidate_contributions %>%
  #joining candidates, committees and contributions
  inner_join(committees, by = "cmte_id") %>%
  #select relevant columns
  select(cmte_id, cmte_name, cmte_party_affiliation, transaction_type, name, transaction_dt, transaction_amt, cand_name, cand_party_affiliation, cand_ici, state)
```

```{r}
#set the desired values of transaction_type
#transaction <- c("24A", "24E")
transaction <- c("24E") #look for transactions advocating election of a candidate
committee_to_cand <- committee_to_cand %>%
  #filter transaction_type to values in 'transaction' i.e either "24A" or "24E"
  filter(transaction_type %in% transaction) %>%
  mutate(transaction_dt = mdy(transaction_dt))
```

```{r}
#filter date to year 2012 i.e transactions between Jan 1st, 2012 to Dec 31st, 2012
committee_trans_2012 <- committee_to_cand %>%
  filter(transaction_dt >= "2012-01-01" & transaction_dt <= "2012-12-31")
head(committee_trans_2012)
unique(committee_trans_2012$cmte_name)[c(1,2,3,4,5,6)]
```

*Let us begin by looking at the contributions made by a single committee - the "Susan B Anthony List Ince. Candidate Fund"*

```{r}
susanb_cand_fund <- committee_trans_2012 %>%
  #filtered for a certain committee
  filter(cmte_name == "SUSAN B ANTHONY LIST INC. CANDIDATE FUND") %>%
  #grouped by whether candidate is incumbent, challenger or 
  group_by(cmte_name, cand_ici, state) %>%
  summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))

committee_summary <- function(data_trans, cmt_name)
  data_trans %>%
    #filtered for a certain committee
    filter(cmte_name == cmt_name) %>%
    #grouped by whether candidate is incumbent, challenger or 
    group_by(cmte_name, cand_ici) %>%
    summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))
```

```{r}
#global variables that can be accessed by all functions and plots
col1 <- "#EBDF8E"
col2 <- "#EFC1CF"
col3 <- "#83C6CC"
font <- "sans"
```

From the graph below we see that in the states of Vermont  & Florida most contributions were given to the incumbent in comparison to their challenger and open seat. 

```{r , echo=FALSE}
#insert a fill inorder to see the transactions to different candidate types
ggplot(susanb_cand_fund, aes(x = state, y = cont_to_ici,
                             fill = cand_ici))+
  #flip the x and y coordinates
  coord_flip()+
  ggtitle("Susan B Anthony List Inc. Candidate Fund")+
  geom_bar(stat = "identity") + 
  #view values on the geom_bar
  geom_text(aes(label=cont_to_ici),position=position_stack(vjust=0.5))+
    scale_x_discrete("State") +
  scale_y_continuous("Total '24E' Contributions:\n Transaactions Advocating Election of a Candidate", limits = c(0,10000), breaks = seq(0, 10000, 5000), labels = c("$0", "$5,000", "$10,000")) +
  scale_fill_manual(values = c(col1, col2, col3),
                    name = "Candidate\nStatus",
                    breaks = c("C", "I", "O"),
                    labels = c("Challenger", "Incubent", "Open Office"))+
  theme(plot.title = element_text(lineheight=.8, face="bold", family = font),
        axis.title.x = element_text(size = 9, face = "bold", family = font),
        axis.title.y = element_text(size = 9, face = "bold", family = font),
        #place legend/key in the bottom
        legend.position = "bottom",
        legend.title = element_text(size = 9, face = "bold"))
```

```{r}
plot_cmt_summary <- function(cmt_summary,cmt_name)
  ggplot(cmt_summary, aes(x = cmte_name, y = cont_to_ici,
                          fill = cand_ici))+
  coord_flip() +
  geom_bar(stat = "identity") + 
  scale_x_discrete("Committee", labels = c(cmt_name)) +
  scale_y_continuous("Total Contributions", limits = c(0,sum(cmt_summary$cont_to_ici)), breaks = seq(0,sum(cmt_summary$cont_to_ici), sum(cmt_summary$cont_to_ici)/5), labels = c("0%","20%","40%","60%","80%","100%") ) +
  scale_fill_manual(values = c("#999999", "#E69F00", "#56B4E9"),
                    name = "Candidate\nStatus",
                    breaks = c("C", "I", "O"),
                    labels = c("Challenger", "Incubent", "Open Office")) 

#  theme(plot.title = element_text(lineheight=.8, face="bold", family = font),
#       axis.title.x = element_text(size = 9, face = "bold", family = font),
#      axis.title.y = element_text(size = 9, face = "bold", family = font),
#     legend.position = "bottom",
#    legend.title = element_text(size = 9, face = "bold"))
#  guides(fill = guide_legend("Candidate
#Status"))
```


```{r  fig.width=5, fig.height=0.75, echo=FALSE}
susanb_cand_fund <- committee_summary(committee_trans_2012,"SUSAN B ANTHONY LIST INC. CANDIDATE FUND")
plot_cmt_summary(susanb_cand_fund,"Susan Anthony")
```

```{r , echo=FALSE}
amer_cand_fund <- committee_summary(committee_trans_2012,"CLUB FOR GROWTH PAC"  )
plot_cmt_summary(amer_cand_fund,"CLUB FOR GROWTH PAC"  )
```


```{r}
#function:
contributions <- function(com_name) {
  table <- committee_trans_2012 %>%
  #filtered for a certain committee
  filter(cmte_name == com_name) %>%
  #grouped by whether candidate is incumbent, challenger or 
  group_by(cmte_name, cand_ici, state) %>%
  summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))
head(table)

ggplot(table, aes(x = state, y = cont_to_ici,
                             fill = cand_ici))+
  coord_flip()+
  ggtitle(com_name)+
  geom_bar(stat = "identity") + 
  geom_text(aes(label=cont_to_ici),position=position_stack(vjust=0.5))+
  scale_x_discrete("State") +
  scale_y_continuous("Total '24E' Contributions:\n Transaactions Advocating Election of a Candidate", limits = c(0,10000), breaks = seq(0, 10000, 5000), labels = c("$0", "$5,000", "$10,000")) +
  scale_fill_manual(values = c(col1, col2, col3),
                    name = "Candidate\nStatus",
                    breaks = c("C", "I", "O"),
                    labels = c("Challenger", "Incubent", "Open Office"))+
  theme(plot.title = element_text(lineheight=.8, face="bold", family = font),
        axis.title.x = element_text(size = 9, face = "bold", family = font),
        axis.title.y = element_text(size = 9, face = "bold", family = font),
        legend.position = "bottom",
        legend.title = element_text(size = 9, face = "bold"))
}
```


```{r}
com_name <- c("SUSAN B ANTHONY LIST INC. CANDIDATE FUND", "CLUB FOR GROWTH PAC", "SUSAN B ANTHONY LIST INC", "NATIONAL RIFLE ASSOCIATION OF AMERICA POLITICAL VICTORY FUND")
```

```{r, warning = FALSE}
lapply(com_name, FUN = contributions)
```
In order to see the '24E' contributions made by other committees, once can add the committee name to the com_name list.
```{r}
#create a list with variables to be used in the 'contributions' function

com_name <- c("SUSAN B ANTHONY LIST INC. CANDIDATE FUND", "CLUB FOR GROWTH PAC", "SUSAN B ANTHONY LIST INC", "NATIONAL RIFLE ASSOCIATION OF AMERICA POLITICAL VICTORY FUND")
```
```{r}
#apply all values in the list 'com_name' to the contributions function
lapply(com_name, FUN = contributions)
```


**Analysis**

The results are inconsistent with the research that we have found thus far.However,it is important to realize that we have only focused on three committees. Amongst the three committees that we have looked at so far only the  NRA has donated more money to candidates who were running for an Open office in 2012. Also in the Susan B. Anthony committee has in Florida and Virgina donated the most money to incumbents. However, if looking at all the states it gave most of its money to the challengers. 

**Limitations and Future Work**
We have only looked at three committees it will be important to look at all the committees and see if this inconsistency matches or does it hold up with the reasearch that we have found saying to 





