---
title: "Mini-Project 2"
author: "Sneha, Pema, Tenzin"
date: "October 31, 2017"
output: html_document
---

##The Incumbent Advantage


Before we begin let us define some basic terms:

**Who is the ‘incumbent’?**

*The ‘incumbent’ is the current holder of an office or a political post.*

**Who is the ‘challenger’?**

*The ‘challenger’ is the person who wants to hold the office or political post next, i.e the person challenging the incumbent.* 

*For example, in the 2012 U.S. presidential elections, Mitt Romney was challenger to Barack Obama’s incumbency.*

**What is an 'Open Seat?**

*An election with no incumbent runner is called an open seat.*


**What is the Incumbent Advantage?**
 
As the incumbent, one holds many advantages over the challenger. Some of these advantages come in the form of:

- Funding

- Name recognition

- National attention

- Fundraising and Campaign Bases

- Control over the instruments of govt.

- Successful Campaign Experience

- Presumption of success and voters

For our project we are looking at whether incumbents held a monetary advantage over their challengers in the U.S. elections for 2012. In order to do this we will take a look at monetary contributions by committees to challengers, incumbents, and open office by state. More specifically we will be looking at transactions of type '24E', i.e. transactions made to advocate the election of a candidate.

```{r, include=FALSE, message=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
```
```{r eval=FALSE}
head(house_elections)
head(committees)
head(contributions)
```


```{r}
candidate_contributions <- candidates %>%
  #joining candidates and contributions by "cand_id"
  inner_join(contributions, by = "cand_id") %>%
  #selecting the relevant variables
  select(cand_id, cand_name, cand_ici, cand_election_yr, cand_party_affiliation, cmte_id, transaction_type, name, transaction_dt, transaction_amt, state)
```

```{r}
committee_to_cand <- candidate_contributions %>%
  #joining candidates, committees and contributions
  inner_join(committees, by = "cmte_id") %>%
  #select relevant columns
  select(cmte_id, cmte_name, cmte_party_affiliation, transaction_type, name, transaction_dt, transaction_amt, cand_name, cand_party_affiliation, cand_ici, state)
```

```{r}
#set the desired values of transaction_type
#transaction <- c("24A", "24E")
transaction <- c("24E") #look for transactions advocating election of a candidate
committee_to_cand <- committee_to_cand %>%
  #filter transaction_type to values in 'transaction' i.e either "24A" or "24E"
  filter(transaction_type %in% transaction) %>%
  mutate(transaction_dt = mdy(transaction_dt))
```

```{r}
#filter date to year 2012 i.e transactions between Jan 1st, 2012 to Dec 31st, 2012
committee_trans_2012 <- committee_to_cand %>%
  filter(transaction_dt >= "2012-01-01" & transaction_dt <= "2012-12-31")
head(committee_trans_2012)
unique(committee_trans_2012$cmte_name)[c(1,2,3,4,5,6,7,8,9)]
```

*Let us begin by looking at the contributions made by a single committee - the "Susan B Anthony List Ince. Candidate Fund"*

```{r}
susanb_cand_fund <- committee_trans_2012 %>%
  #filtered for a certain committee
  filter(cmte_name == "SUSAN B ANTHONY LIST INC. CANDIDATE FUND") %>%
  #grouped by whether candidate is incumbent, challenger or 
  group_by(cmte_name, cand_ici, state) %>%
  summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))

committee_summary <- function(data_trans, cmt_name)
  data_trans %>%
    #filtered for a certain committee
    filter(cmte_name == cmt_name) %>%
    #grouped by whether candidate is incumbent, challenger or open seat.
    group_by(cmte_name, cand_ici) %>%
    summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))
```

```{r}
#global variables that can be accessed by all functions and plots
col1 <- "#EBDF8E"
col2 <- "#EFC1CF"
col3 <- "#83C6CC"
font <- "sans"
```

From the graph below we see that in the states of Vermont  & Florida most contributions were given to the incumbent in comparison to their challenger and open seat. 

```{r , echo=FALSE}
#insert a fill inorder to see the transactions to different candidate types
ggplot(susanb_cand_fund, aes(x = state, y = cont_to_ici,
                             fill = cand_ici))+
  #flip the x and y coordinates
  coord_flip()+
  ggtitle("Susan B Anthony List Inc. Candidate Fund")+
  geom_bar(stat = "identity") + 
  #view values on the geom_bar
  geom_text(aes(label=cont_to_ici),position="stack", angle = -90, size = 3)+
    scale_x_discrete("State") +
  scale_y_continuous("Total '24E' Contributions:\n Transaactions Advocating Election of a Candidate", limits = c(0,10000), breaks = seq(0, 10000, 5000), labels = c("$0", "$5,000", "$10,000")) +
  scale_fill_manual(values = c(col1, col2, col3),
                    name = "Candidate\nStatus",
                    breaks = c("C", "I", "O"),
                    labels = c("Challenger", "Incubent", "Open Office"))+
  theme(plot.title = element_text(lineheight=.8, face="bold", family = font),
        axis.title.x = element_text(size = 9, face = "bold", family = font),
        axis.title.y = element_text(size = 9, face = "bold", family = font),
        #place legend/key in the bottom
        legend.position = "bottom",
        legend.title = element_text(size = 9, face = "bold"))
```

```{r}
#create the functions 'contributions' that takes in the variable com_name
contributions <- function(com_name) {
  #create a tbl_df to be used to create a ggplot
  table <- committee_trans_2012 %>%
  #filtered for a certain committee by the committee name
  filter(cmte_name == com_name) %>%
  #grouped by whether candidate is incumbent, challenger or open seat
  group_by(cmte_name, cand_ici, state) %>%
  summarise(cont_to_ici = sum(transaction_amt, na.rm = TRUE))
head(table)

#create a ggplot using the tbl_df created above
ggplot(table, aes(x = state, y = cont_to_ici,
                             fill = cand_ici))+
  #flip the coordinates for a better visualisation of the data
  coord_flip()+
  #set the title of the ggplot as the committee name
  ggtitle(com_name)+
  geom_bar(stat = "identity") + 
  #include labels of the '24E' amount contributed to each candidacy
  geom_text(aes(label=cont_to_ici),position="stack", angle = -90, size = 3)+
  #x-axis label
  scale_x_discrete("State") +
  #y-axis label with breaks
  scale_y_continuous("Total '24E' Contributions:\n Transaactions Advocating Election of a Candidate", limits = c(0,10000), breaks = seq(0, 10000, 5000), labels = c("$0", "$5,000", "$10,000")) +
  #set color for each geom_bar and change the legend labels
  scale_fill_manual(values = c(col1, col2, col3),
                    name = "Candidate\nStatus",
                    breaks = c("C", "I", "O"),
                    labels = c("Challenger", "Incubent", "Open Office"))+
  #customise the text and labels in the ggplot
  theme(plot.title = element_text(lineheight=.8, face="bold", family = font),
        axis.title.x = element_text(size = 9, face = "bold", family = font),
        axis.title.y = element_text(size = 9, face = "bold", family = font),
        #place the legend on the bottom
        legend.position = "bottom",
        legend.title = element_text(size = 9, face = "bold"))
}
```


In order to see the '24E' contributions made by other committees, once can add the committee name to the com_name list.

```{r}
#create a list with variables to be used in the 'contributions' function

com_name <- c("SUSAN B ANTHONY LIST INC. CANDIDATE FUND", "CLUB FOR GROWTH PAC", "NATIONAL RIFLE ASSOCIATION OF AMERICA POLITICAL VICTORY FUND", "NATIONAL RIGHT TO LIFE POLITICAL ACTION COMMITTEE", "AMERICANS FOR RESPONSIBLE LEADERSHIP")
```

```{r, warning=FALSE}
#apply all values in the list 'com_name' to the contributions function
lapply(com_name, FUN = contributions)
```


**Analysis**

The results are consistent (for 3 of the 4 committees we have looked at) with the research that we have found thus far. However,it is important to realize that we have only focused on three committees. Amongst the 4 committees that we have looked at so far only the NRA has donated more money to candidates who were running for an Open office in 2012. Also in the Susan B. Anthony committee has in Florida and Virgina donated the most money to incumbents. However, if looking at all the states it gave most of its money to the challengers. 

**Limitations and Future Work**

We have only looked at three committees it will be important to look at all the committees to see if this inconsistency that we have seen thus far holds. Looking at more committees we will be able to see if Incumbents have truly have a monetary advantage against or candidate. another limitation of our graph is that have shown the percentages only the total monetary value. It would be nice to calculate the percentages given to each candidate rather than the amount of money given. In the future, it would interesting to calculate the percentages of contribution given by each committee. Also it would be interesting to see the percentage of Incumbents winning in each state maybe seeing if there are some states that elect more incumbents than others. 

- we need to make labels clearer
- look at different years and transaction types





